### 2024-10-10 authorshipAndContributions.tex

\authorrules{Charles Babbage (Corresponding Author)}
{Computer Scientist, Inventor}
{https://orcid.org/0000-0000-0000-0000}
{charlesbabbage@example.com}
{Conceptualization; Methodology; Writing – Original Draft Preparation}
% 
\authorrules{Ada Lovelace}
{Mathematician, Writer}
{} 
{adalovelace@example.com}
{Literature Review, Data Analysis, Writing – Review \& Editing}
% 
\authorrules{Pierre Curie}
{Physicist}
{https://orcid.org/0000-0000-0000-0000}
{pierrecurie@example.com}
{Methodology, Data Analysis, Writing – Review \& Editing}
% 
\authorrules{Marie Curie}
{Physicist, Chemist}
{https://orcid.org/0000-0000-0000-0000}
{mariecurie@example.com}
{Conceptualization, Methodology, Data Analysis, Writing – Review \& Editing}
% 
\authorrules{Grace Hopper}
{Computer Scientist}
{https://orcid.org/0000-0000-0000-0000}
{gracehopper@example.com}
{Conceptualization, Methodology, Writing – Original Draft Preparation}
% 
\authorrules{Santos Dumont}
{Inventor, Aviator}
{https://orcid.org/0000-0000-0000-0000}
{santosdumont@example.com}
{Conceptualization, Methodology, Writing – Original Draft Preparation}
%
\authorrules{Nikola Tesla}
{Inventor, Electrical Engineer}
{} 
{nikolatesla@example.com}
{Methodology, Data Analysis, Writing – Review \& Editing}
%